---
title: "STAT40830 – Advance Data Programming with R"
author: "Charukeshi Senthamarai Kannan – 24213384"
format:
  revealjs:
    theme: "serif"
    code-visibility: show
    slide-number: true
execute:
  echo: true
  warning: false
  message: false
fontsize: 15pt
---

## Assignment 1

#### Introduction:

In this presentation, we analyze the socio-economic, environmental, and educational trends of three countries: India, Brazil, and Germany using "data.table" .

#### Question 1

Loading the Libraries

```{r}
library(data.table)
library(ggplot2)
```

Loading and cleaning the datasets

```{r}
col_names <- c("Country Name", "Country ISO3", "Year", 
               "Indicator_Name", "Indicator_Code", "Value")

india <- fread("india.csv", skip = 1, col.names = col_names)
brazil <- fread("brazil.csv", skip = 1, col.names = col_names)
germany <- fread("germany.csv", skip = 1, col.names = col_names)
```

Defining a cleaning function to assign correct classes and to add country name

```{r}
clean_data <- function(dt, country_name) {
  dt[, Year := as.integer(Year)]                 
  dt[, Value := as.numeric(Value)]               
  dt[, Country := country_name]                  
  return(dt)
}
```

Applying the cleaning function to each dataset

```{r}
india <- clean_data(india, "India")
brazil <- clean_data(brazil, "Brazil")
germany <- clean_data(germany, "Germany")
```

------------------------------------------------------------------------

#### Question 2:

Combining all the three cleaned datasets into one using rbindlist

```{r}
comb_data <- rbindlist(list(india, brazil, germany), use.names = TRUE, fill = TRUE)
```

------------------------------------------------------------------------

#### Question 3:

```{r}
library(skimr)
library(dplyr)
```

checking the structure and the summary of the merged data

```{r}
str(comb_data)
```

---------------------------------------------------------------------------------------

Summary of the combined data

```{r}
summary(comb_data)
```

---------------------------------------------------------------------

Using skimr for a detailed overview (type, missing, distribution)

```{r}
skim(comb_data)
```
-----------------------------------------------------------------------

Counting the number of unique indicators

```{r}
comb_data[, .(Num_Indicators = uniqueN(Indicator_Name))]
```

------------------------------------------------------------------------

Finding the top 10 most frequent indicators

```{r}
comb_data[, .N, by = Indicator_Name][order(-N)][1:10]
```

------------------------------------------------------------------------

Counting the number of rows per country

```{r}
comb_data[, .N, by = Country]
```

Counting the number of years per country

```{r}
comb_data[, .(NumYears = uniqueN(Year)), by = Country]
```

------------------------------------------------------------------------

[**Summary of the outputs from data exploration:**]{.underline}

-   The combined dataset contains 252,741 rows and 4,551 unique indicators, covering India, Brazil, and Germany across 65 years (1960–2024).

<!-- -->

-   Some indicators appear consistently, such as Net migration, Life expectancy, and Agricultural land, indicating common global metrics.

------------------------------------------------------------------------

#### Question 4:

[**Analysis 1:** Average value of each indicator per country]{.underline}

```{r}
avg_indicator_by_country <- comb_data[
  , .(Avg_Value = mean(Value, na.rm = TRUE)), 
  keyby = .(Country, Indicator_Name)
]

head(avg_indicator_by_country)
```

The average value calculation by indicator and country shows Brazil's indicators listed first due to "keyby" sorting alphabetically by Country and then by Indicator Name.

------------------------------------------------------------------------

[**Analysis 2:** Trend of a specific indicator (Life expectancy at birth, female (years))]{.underline}

```{r}
life_expectancy_female <- comb_data[
  Indicator_Name == "Life expectancy at birth, female (years)"
]

avg_life_expectancy_female <- life_expectancy_female[
  , .(Avg_LifeExp = mean(Value, na.rm = TRUE)), 
  by = .(Year, Country)
]

avg_life_expectancy_female

```

-   Across the years, Germany consistently maintains the highest female life expectancy, followed by Brazil, while India shows steady improvement from 45 to over 73 years.

-   All three countries exhibit a long-term upward trend, with India experiencing the most significant relative growth since 1960.

------------------------------------------------------------------------

[**Analysis 3:** Identifying the Top 5 Indicators with Highest Average Value per Country]{.underline}

```{r}
top_indicators_by_country <- comb_data[
  , .(Avg_Value = mean(Value, na.rm = TRUE)), 
  by = .(Country, Indicator_Name)
][order(-Avg_Value), .SD[1:5], by = Country]

top_indicators_by_country

```

------------------------------------------------------------------------

[**Analysis 4:** Trends in Agricultural Land]{.underline}

```{r}
agri_land <- comb_data[
  Indicator_Name == "Agricultural land (% of land area)"
]

agri_land_trend <- agri_land[
  , .(Avg_Land = mean(Value, na.rm = TRUE)), 
  by = .(Year, Country)
]

head(agri_land_trend)
```

------------------------------------------------------------------------

#### Question 5:

[**Life Expectancy at Birth (Female) Over Time for All Countries**]{.underline}

```{r}
ggplot(avg_life_expectancy_female, aes(x = Year, y = Avg_LifeExp, color = Country)) +
  geom_line(size = 1.1) +
  labs(
    title = "Female Life Expectancy at Birth (1960–2023)",
    x = "Year",
    y = "Average Life Expectancy (Years)"
  ) +
  theme_minimal()
```

**Observation:**

-   Germany consistently leads in female life expectancy, followed by Brazil and India, all showing steady improvements over time.

- Female life expectancy in Europe is comparatively high when compared to other place 

-   A noticeable dip around 2020–2021 across all countries likely reflects the impact of the COVID-19 pandemic.

------------------------------------------------------------------------

[**Number of Indicators Reported per Year by Country**]{.underline}

```{r}
indicator_count <- comb_data[
  , .(Num_Indicators = uniqueN(Indicator_Name)), 
  by = .(Year, Country)
]
ggplot(indicator_count, aes(x = Year, y = Num_Indicators, fill = Country)) +
  geom_area(position = "identity", alpha = 0.5) +
  labs(
    title = "Number of Unique Indicators Reported per Year",
    x = "Year",
    y = "Indicator Count"
  ) +
  theme_minimal()
```

**Observation:**

-   The number of reported indicators increased steadily for all countries until around 2015, with India consistently reporting the most.

-   Post-2015, Brazil shows a relative surge in reporting, while a decline in all countries is visible after 2020, possibly due to data lags or disruptions.

------------------------------------------------------------------------
